Spin canonical divisors of log stacky curves

### 2015-06-20 stacky_write_up/spin_cring.tex
proving b

\newcommand \subhalf[1]{\frac{e_{#1} - 1}{2e_{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}

@@

\section{Introduction}
This is our project.
Example, O'Dorney's paper \cite{dorney:canonical}.
We consider divisors of the form $D = \sum_{i = 1}^{r} \subhalf{i}$.

@@

\begin{lem}
\label{sat_three_induction}
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted  log stacky curves $R \subset R'$ be the corresponding inclusion of canonical rings. Let $R_{D'} = R' = k[x]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $D = \subhalf i$. There are generators $y_3,y_4,y_5$ of $R_{D+D'}= R$ over $R'$ so that $y_i$ lies in degree $i$ and has a pole of order $1$. \todo{Include the order}
Let $x_3,x_4,x_5$ be generators of $k[x]$ lying in degrees 3,4,5 respectively. \todo{Make sure to state some requirement on the ordering of the $x_i$, analogously to 8.4.1, but I don't see why this is important.} Then, letting $x_j$ range over all $x$ variables
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y_3,\ldots, y_e,z_4,z_5] + \\
	\langle y_i y_j \mid 1 \leq i < j-3 \leq e \rangle +\\
	\langle y_i x_j \mid i > 3 \rangle +\\
	\langle z_i y_j \mid j > 3 \rangle  +\\
	\langle z_i x_j \mid 4 \leq i \leq 5\rangle + \\
	\langle z_iz_k \mid 4 \leq i \leq j \leq 5 \rangle 
\end{align*}

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx',L'),(\sx,L)$ be tame, separably rooted log stacky spin curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose there is a birational map $\sx \rightarrow \sx'$, and let $J \subset \{1,
\ldots, r\}$. Suppose $e_i'+ 2\chi_J(i) = e_i$ where
$$
\chi_J(i) = \begin{cases}
	1, &\text{ if }i \in J\\
	0, &\text{ otherwise. } 
\end{cases}$$
Let $R'$ be the canonical ring associated to $\sx'$. Say $(\sx',J)$, {\bf admissible} if the following three conditions hold.
\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} \in R'$ so that $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
\end{align*}
so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
\item[(Ad-ii)] For all $i \in J$ and $z \neq y_{i,e_i'}$ a generator of $R'$ we have
\begin{align*}
	\frac{- ord_{Q_i}(z)}{\deg z} < \frac{e_i'-1}{2(e_i')}.
\end{align*}
\item[(Ad-iii)] For all $i \in J$ we have
\begin{align*}
	\deg \lfloor e_i L \rfloor \geq \# S(i)
\end{align*}
where
\begin{align*}
	\eta(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
\end{enumerate}
\end{defn}

\begin{lem}
\label{lem:admissible_inequality}
Condition (Ad-ii) of Definition ~\ref{defn:admissible} implies the stronger condition that
\begin{align*}
	-\ord_{Q_i}(z) \leq \deg(z) \frac{e_i'-1}{2e_i'}-\frac{1}{e_i'}
\end{align*}
\end{lem}
\begin{proof}
We know by (Ad-ii) that
\begin{align*}
	-\ord_{Q_i}(z) < \deg(z) \frac{e_i'-1}{2e_i'}
\end{align*}
If we write $\frac{\alpha}{\beta} = \deg(z) \frac{e_i'-1}{2e_i'}$ as a fraction in lowest terms, then we see $\beta \mid e_i'$ since $e_i'-1$ is even. Therefore, since $-\ord_{Q_i}(z)$ is an integer, we must have
\begin{align*}
	-\ord_{Q_i}(z) \leq \deg(z) \frac{e_i'-1}{2e_i'} - \frac{1}{\beta} \leq \deg(z) \frac{e_i'-1}{2e_i'} - \frac{1}{e_i'}.
\end{align*}
\end{proof}

\begin{thm}
\label{thm:ramification_order_induction}
Suppose $(\sx',L')$ is a tame separably rooted log stacky spin curve is admissible with a generators $y_{i,e_i'} \in R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in H^0(\sx, e_i(K_\sx))$ so that
	\begin{align*}
	-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
\end{align*}
and
\begin{align*}
	\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \frac{e_j'-1}{2e_j'}-\frac{1}{\deg(y_{i,e_i'})e_j'}
\end{align*}
for all $j \in J$ with $j \neq i$.
\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0, b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally generate $R$ over $R'$.
\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with block order. Let $R = k[y,x]/I.$ Then,
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z \neq y_{i,e_i},y_{i,e_i'} \rangle 
\end{align*}
	as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by quadratics and that for all $i \in J$, we have $e_i > \deg z$ for any generator $z$ of $R'.$ Then any set of minimal generators for $I'$ together with any set of relations as in the previous part \todo{reference previous part as part c} minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}
\begin{proof}
Let $(\sx,L)$ and $(\sx',L')$ be the two log stacky spin curves in question. Define
\begin{align*}
	S(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}

>>>>

\newcommand \subhalf[1]{\frac{{#1} - 1}{2{#1}}}
\newcommand{\se}[1]{\section*{Problem #1}}

@@

\section{Introduction}
This is our project.
Example, O'Dorney's paper \cite{dorney:canonical}.
We consider divisors of the form $D = \sum_{i = 1}^{r} \subhalf{e_i}$.

@@

\begin{lem}
\label{sat_three_induction}
Let $\sx \rightarrow \sx'$ be a birational map of tame, separably rooted  log stacky curves $R \subset R'$ be the corresponding inclusion of canonical rings. Let $R_{D'} = R' = k[x]/I'$ satisfy $sat(Eff(D')) = 3.$ Let $D = \subhalf {e_i}$. There are generators $y_3,y_4,y_5$ of $R_{D+D'}= R$ over $R'$ so that $y_i$ lies in degree $i$ and has a pole of order $1$. \todo{Include the order}
Let $x_3,x_4,x_5$ be generators of $k[x]$ lying in degrees 3,4,5 respectively. \todo{Make sure to state some requirement on the ordering of the $x_i$, analogously to 8.4.1, but I don't see why this is important.} Then, letting $x_j$ range over all $x$ variables
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y_3,\ldots, y_e,z_4,z_5] + \\
	\langle y_i y_j \mid 1 \leq i < j-3 \leq e \rangle +\\
	\langle y_i x_j \mid i > 3 \rangle +\\
	\langle z_i y_j \mid j > 3 \rangle  +\\
	\langle z_i x_j \mid 4 \leq i \leq 5\rangle + \\
	\langle z_iz_k \mid 4 \leq i \leq j \leq 5 \rangle 
\end{align*}

@@

\begin{defn}
\label{defn:admissible}
Let $(\sx',L'),(\sx,L)$ be tame, separably rooted log stacky spin curves, both ramified over ramified at $Q_1,\ldots, Q_r$. Suppose there is a birational map $\sx \rightarrow \sx'$, and let $J \subset \{1,
\ldots, r\}$. Suppose $e_i'+ 2\chi_J(i) = e_i$ where
$$
\chi_J(i) = \begin{cases}
	1, &\text{ if }i \in J\\
	0, &\text{ otherwise. } 
\end{cases}$$
Let $R'$ be the canonical ring associated to $\sx'$. Say $(\sx',J)$, {\bf admissible} if the following three conditions hold.
\begin{enumerate}
	\item[(Ad-i)] Suppose for all $i \in J$, there exists $y_{i,e_i'} \in R'$ so that $R'$ admits a presentation
\begin{align*}
	R' \cong \left( k[x] \otimes k[y_{i,e_i'}]_{i \in J} \right)/I'
\end{align*}
so that for all $i \in J$ we have $\deg y_{e_i'} = e_i'$ and $-ord_{Q_i}(y_{i,e_i'}) = \frac{e_i'-1}{2}$.
\item[(Ad-ii)] For all $i \in J$ and $z \neq y_{i,e_i'}$ a generator of $R'$ we have
\begin{align*}
	\frac{- ord_{Q_i}(z)}{\deg z} < \subhalf {e_i'}.
\end{align*}
\item[(Ad-iii)] For all $i \in J$ we have
\begin{align*}
	\deg \lfloor e_i L \rfloor \geq \# S(i)
\end{align*}
where
\begin{align*}
	\eta(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
\end{enumerate}
\end{defn}

\begin{lem}
\label{lem:admissible_inequality}
Condition (Ad-ii) of Definition ~\ref{defn:admissible} implies the stronger condition that
\begin{align*}
	-\ord_{Q_i}(z) \leq \deg(z) \subhalf{e_i'}-\frac{1}{e_i'}
\end{align*}
\end{lem}
\begin{proof}
We know by (Ad-ii) that
\begin{align*}
	-\ord_{Q_i}(z) < \deg(z) \subhalf{e_i'}
\end{align*}
If we write $\frac{\alpha}{\beta} = \deg(z) \frac{e_i'-1}{2e_i'}$ as a fraction in lowest terms, then we see $\beta \mid e_i'$ since $e_i'-1$ is even. Therefore, since $-\ord_{Q_i}(z)$ is an integer, we must have
\begin{align*}
	-\ord_{Q_i}(z) \leq \deg(z) \subhalf{e_i'}- \frac{1}{\beta} \leq \deg(z) \subhalf{e_i'}- \frac{1}{e_i'}.
\end{align*}
\end{proof}

\begin{thm}
\label{thm:ramification_order_induction}
Suppose $(\sx',L')$ is a tame separably rooted log stacky spin curve is admissible with a generators $y_{i,e_i'} \in R' = R_{L'}$ as in (Ad-i) of Definition ~\ref{defn:admissible}. Then the following are true:
\begin{enumerate}
	\item[(a)] For all $i \in J$, there exists $y_{i,e_i} \in H^0(\sx, e_i(K_\sx))$ so that
	\begin{align*}
	-ord_{Q_i}(y_{i,e_i}) = \frac{e_i-1}{2}
\end{align*}
and
\begin{align*}
	\frac{-\ord_{Q_j}(y_{i,e_i'})}{y_{i,e_i}} \leq \subhalf{e_j'}-\frac{1}{\deg(y_{i,e_i'})e_j'}
\end{align*}
for all $j \in J$ with $j \neq i$.
\item[(b)] The elements $y_{i,e_i'}^ay_{i,e_i}^b,$ with $a \geq 0, b > 0$ span $R$ over $R'$ and the elements $y_{i,e_i}$ minimally generate $R$ over $R'$.
\item[(c)] Equip $k[y] = k[y_{i,e_i}]_{i \in J}$ and $k[x]$ with any graded monomial order and $k[y,x] = k[y] \otimes k[x]$ with block order. Let $R = k[y,x]/I.$ Then,
\begin{align*}
	in_\prec(I) = in_\prec(I')k[x,y] + \langle y_{i,e_i}z \mid z \neq y_{i,e_i},y_{i,e_i'} \rangle 
\end{align*}
	as $z$ ranges over all generators of $R$.
	\item[(d)] Suppose $in_\prec(I')$ is minimally generated by quadratics and that for all $i \in J$, we have $e_i > \deg z$ for any generator $z$ of $R'.$ Then any set of minimal generators for $I'$ together with any set of relations as in the previous part \todo{reference previous part as part c} minimally generate $I$.
	\item[(e)] For any $W \subset J,$ the set $(\sx,W)$ is admissible.
\end{enumerate}
\end{thm}
\begin{proof}
Let $(\sx,L)$ and $(\sx',L')$ be the two log stacky spin curves in question, with coarse spaces $X = X'$. Define
\begin{align*}
	S(i) = \{j \in J : j \neq i \text{ and }e_j' \mid e_i\}
\end{align*}
Define
\begin{align*}
	E_i = \sum_{j \in S(i)}^{}Q_j
\end{align*}
Note that
\begin{align*}
	\lfloor e_i L' \rfloor + Q_i \leq \lfloor e_i L \rfloor 
\end{align*}
and so we obtain an inclusion
\begin{align*}
	H^0(\sx',e_iL'-E_i + Q_j) \rightarrow H^0(\sx,e_iL - E_i) \subset H^0(\sx,e_iL).
\end{align*}
The assumption (Ad-iii) implies
\begin{align*}
	\deg \left( e_i L' - E_i \right) \geq 0
\end{align*}
implying that $H^0(\sx',e_iL'-E_i + Q_i)$ is base point free by Riemann Roch, and so a general element
\begin{align*}
	y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)
\end{align*}
satisfies
\begin{align*}
	-\ord_{Q_i}(y_{i,e_i}) = \lfloor e_i \subhalf {e_i'} \rfloor +1 = \frac{e_i - 1}{2}
\end{align*}
Therefore, $y_{i,e_i}$ satisfies the first part of the claim of $(a)$.

We next show $y_{i,e_i}$ also satisfies the second part of the claim of $(a),$ by considering seperately the cases of whether $j \in S(i),$ and $j \notin S(i)$.

If $j \in S(i)$ then, since $E_i \geq Q_j$, and $y_{i,e_i} \in H^0(\sx',e_iL'-E_i + Q_i)$,
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq e_i\subhalf {e_j'} - 1 \leq e_i \subhalf{e_j'} - \frac{1}{e_j'}.
\end{align*}

If instead $j \notin S(i)$, then since $e_j' \nmid e_i$, we know $e_i\subhalf{e_j'} \notin \BZ$, so
\begin{align*}
	-\ord_{Q_j}(y_{i,e_i}) \leq \lfloor  e_i\subhalf{e_j'} \rfloor \leq e_i\subhalf{e_j'} - \frac{1}{e_j'},
\end{align*}
completing the proof of (a).

Next, define $R_0 = R'$ and inductively define 
$$
R_i = \begin{cases}
	R_{i-1} &\text{ if }i \notin J\\
	R_{i-1}[y_{i,e_i}] &\text{ otherwise.} 
\end{cases}$$
To prove (b), it suffices to show that elements of the form $y_{i,e_i'}^ay_{i,e_i}^b$ with $a \geq 0, b > 0$ are linearly indpendent, and that these elements together with $R_{i-1}$ span $R_i$ as a $k$-vector space. These elements do not lie in $R_{i-1}$ because their order of pole at $Q_i$ is bigger than the order of any element in $R_{i-1}$, and are linearly independent amongst themselves because of injectivity of the linear map
\begin{align*}
	(a,b) \mapsto \left( \deg\left(y_{i,e_i'}^ay_{i,e_i}^b\right),-\ord_{Q_i}\left( y_{i,e_i'}^ay_{i,e_i}^b \right)  \right) = (a,b) \begin{pmatrix}
	e_i' & \subhalf{e_i'} \\
	e_i	 & \subhalf{e_i}
\end{pmatrix} 
\end{align*}
